---
header-includes:
- \input{preamble.tex}
fontsize: 11pt
output: pdf_document
sansfont: Roboto
font:     Roboto
geometry: "left=1in,right=1in,top=0.35in,bottom=0.6in"
urlcolor: #169d62
---

\raggedright

```{r rmarkdown-setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
knitr::opts_chunk$set(message = FALSE)
knitr::opts_chunk$set(warning = FALSE)
knitr::opts_chunk$set(dev = "cairo_pdf")
options(knitr.kable.NA = "")
options(kableExtra.latex.load_packages = FALSE)

library(tidyverse)
library(knitr)
library(kableExtra)
library(sdcoethemes)

set_sdcoe_defaults(style = "print")
```

\sdcoelogo{}

\vspace{5mm}
\sdcoetitle{Declines in College Enrollment Continue Unabated in 2021}
\sdcoeauthors{Shannon Coulter, Director of Research and Evaluation}

\vspace{5mm}
\begin{adjustwidth}{50pt}{50pt}
\lettrine[lines=3, lraise=0.2, nindent=0em, slope=0em]{H}igher education enrollment fell to its lowest levels in 2021. In San Diego, more than 5,250 fewer students opted in to college compared to five years ago. Such a decline poses a serious risk for many students, especially those students who rely on a college degree to become upwardly mobile. And yet, simply enrolling in college is no panacea to upward mobility, students must actually attain a college degree, which is no small feat given that fewer than 25 percent actually do. Because of these historic declines in enrollment, we face a monumental problem: We currently have fewer students in the pipeline to attain a college degree--potentially losing 1,000 degree earners from the Class of 2021 alone--and this problem has the potential to further erode a fundamental purpose of K-12 education, our higher education institutions, and our economy at large.

Decisions about attending college were challenging prior to the pandemic. Students often questioned their readiness for the rigor of college work and the financial costs. The pandemic has layered other concerns that make college-going decisions even more unstable for students. 

There is no doubt that many students need a college degree to attain their professional aspirations and to achieve economic mobility. San Diego also needs degree earners--the relationship between economic growth and educational attainment are undeniable. We need to act on these complex problems now and address them with long terms solutions, and in the short term, we must continue to improve our understanding of our college-going data and use this information to mitigate students' fears and hesitations about attaining a college degree.

We created this annual report for the purpose of increasing our understanding and decisions about college access and completion. Instead of relying on self-reported information about how many students go to college,  we use the National Student Clearinghouse's StudentTracker service, a nationwide source of college enrollment and degree data from over 3,600 colleges and universities. The StudentTracker data provides our community with a reasonably accurate representation of the number and percentage of students who enroll, persist, and complete college annually. This report helps guide our efforts to get our students back into college and keep our economic outlook bright.
\end{adjustwidth}

\contactinfo{}

\newpage{}
\newgeometry{left=1in,right=1in,top=1in,bottom=1.5in}

\sdcoeheadingone{College Enrollment}

In 2021, the National Student Clearinghouse reported 18,915 students enrolled in college right out of high school, a slight decline from the year before. College enrollment patterns in the Clearinghouse may vary from year to year for a variety of reasons including an overall decline in college going, fluctuations in the size of 12th grade cohorts, and the timing of when local high schools submit their graduates to the Clearinghouse. For that reason, our fall report is a preliminary report that is updated in late spring.

The map below locates the nearly 1,900 colleges and universities where students enrolled in 2021. Most students (83%) attended colleges and universities in southern California, while the remaining 2,200 attended institutions in all other 49 states. 

\sdcoefigurenumber{1}
\sdcoefiguretitle{While Most College Students Stay Close to Home, Graduates Enroll in Colleges and Universities Across the Country}

```{r enroll-map}
source("../nsc_21/scripts/09b usa_map.R", chdir = T)
```
\sdcoesource{National Student Clearinghouse StudentTracker Report.}
\sdcoenote{We used geospatial data to identify the location of each college or university where graduates from the Class of 2021 enrolled. We mapped each student to his or her respective college or university, darker dot patterns represent areas where more students attend college.}

\newpage{}
\newgeometry{left=1in,right=1in,top=1in,bottom=1.5in}

\sdcoeheadingtwo{Top Ten Colleges and Universities Attended}

For the first time in eight years more students enrolled in four-year institutions (9,633) compared to 2-year (8,007). Some local community colleges suffered enrollment loses upwards of 25% in 2021, while the University of California-San Diego reported a 26% increase in enrollment compared to the past two years. These variations in enrollment patterns have the greatest impact on lower-income students whose enrollment declines were double other student groups according to the National Student Clearinghouse.

\vspace{3mm}
\sdcoetablenumber{1}
\sdcoetabletitle{More High School Graduates Opted for a Four-Year College Experience in 2021}
\sdcoetablesubtitle{Community colleges still have a big enrollment footprint in San Diego}

```{r echo=FALSE,  fig.width =7 , fig.height =2}
# arbitrary data to demonstrate a table
data <- tribble(~"College Name", ~"Freshman Class", ~"Pell grad share", ~"Net price, middle-income",
  "University of California-Irvine", 5449, 40, "$13k",
  "University of California-Davis", 5063, 31, "$14k",
  "University of California-Santa Barbara", 4597, 31, "$14k",
  "University of California-San Diego", 5218, 28, "$13k",
  "University of California-Los Angeles", 5684, 28, "$13k",
  "University of California-Los Angeles", 5684, 28, "$13k",
  "University of Florida", 6348, 24, "$9k",
  "University of California-Berkeley", 4677, 23, "$13k",
  "Vassar", 662, 22, "$12k",
  "Amherst", 466, 20, "$11k",
  "Pomona", 396, 18, "$9k",
  )

kable(data, format = "latex", booktabs = TRUE,
      col.names = c(" ", "Freshman Class", "Pell grad share", "Net price, middle-income"),
      align = c("l", "r", "r", "r"),
      linesep = "") %>%
  kable_styling(full_width = TRUE) %>%
  pack_rows("College Name", 1, 10) %>%
  column_spec(1, width = "7cm") %>%
  column_spec(2, width = "4.75cm") %>%
  row_spec(0, bold = TRUE,
           extra_css = "border-top: 0; border-bottom:5px solid red;") %>%
  add_header_above(c(" " , " ", " ", " ", " "))
```
\sdcoesource{National Student Clearinghouse StudentTracker Report.}

\newpage{}
\newgeometry{left=1in,right=1in,top=1in,bottom=1.5in}

\sdcoeheadingtwo{College Enrollment by Cohort}

San Diego's total undergraduate enrollment has declined over 3200 students in the past 2 years. The loss to these students' futures is significant and will greatly impact our community in years to come. We need to re-engage these and all students in the value of a college education.

\sdcoefigurenumber{2}
\sdcoefiguretitle{Early Enrollment Figures Show Another Steep Decline in College-Going}
\vspace{3mm}

```{r enroll-by-cohort}
source("../nsc_21/scripts/019a enroll_cohort.R", chdir = T)
```
\sdcoesource{National Student Clearinghouse StudentTracker Report.}
\sdcoenote{We calculated enrollment numbers for each graduating cohort using data provided by the National Student Clearinghouse. The Clearinghouse data file supplies informatin on whether a student enrolled in college or not with a 93 percent matching rate.}


\newpage{}
\newgeometry{left=1in,right=1in,top=1in,bottom=1.5in}

\sdcoeheadingone{Completion}

While the pandemic has made no discernible difference on college completion rates yet, only a  small number of students actually earn college degrees. About 17% of the Class of 2017 has earned a college degree, while 24% of the Class of 2015 earned a degree. We know that degree completers earn more on average than non-college graduates, and that we need a certain level of degree completers for our economy to thrive. While the percentage of students earning degrees compared to those enrolled may not change much in the coming years, we might have significantly fewer students earning degrees in the future based on our current enrollment levels.

\sdcoefigurenumber{4}
\sdcoefiguretitle{Only 1 in 4 San Diego Students Earn a College Degree}
\sdcoefiguresubtitle{The percentange of students completing degrees in four and six years remains low}
```{r completion-by-cohort}
source("../nsc_21/scripts/012a completion_rates.R", chdir = T)
```
\sdcoesource{National Student Clearinghouse StudentTracker Report.}
\sdcoenote{Colleges report graduation rates many different ways. We calculated graduation rates by cohort or year of high school graduation. The 2015 cohort represents the 6-year graduation cohort and the data point reflects the percentage of students from that cohort who earned a degree.}

\newpage{}
\newgeometry{left=1in,right=1in,top=1in,bottom=1.5in}

\sdcoeheadingtwo{Degrees Earned}

Business Administration degrees top the list as the most popular college choice followed by Psychology and Computer Science. Because earnings vary based on the type of degree students get, several organizations including [College Scorecard](https://collegescorecard.ed.gov/) and [Salary Surfer](https://salarysurfer.cccco.edu/SalarySurfer.aspx) provide valuable information linking college degrees to the median salaries of graduates. Surprisingly the average salary for some Associate's degrees are as much if not more than comparable Bachelor's degrees. Using earnings information can help reduce some students' fears and hesitations about enrolling in college.

\sdcoetablenumber{2}
\sdcoetabletitle{2014 Graduates Earned Over 7500 Degrees}
\sdcoetablesubtitle{Many students earn two-year degrees in same time some students earn four-year degrees}

```{r echo=FALSE,  fig.width =7 , fig.height =2}
# arbitrary data to demonstrate a table
data <- tribble(~"Degree", ~"# Obtained", ~"Time to Completion",
  "Certificate", 313, "3 years 11 months",
  "Associate's", 347, "4 years 1 month",
  "Bachelor's", 3082, "4 years 6 months",
  "Master's and beyond (with Bachelors)", 352, "5 years 11 months",
  "Unreported", 3418, "4 years 6 months",
  "Business Administration", 722, "",
  "Psychology", 503, "",
  "Computer Science", 424, "",
  "Political Science", 256, "",
  "Economics", 254, "",
  "Sociology", 250, "",
  "Communication", 229, "",
  "Mechanical Engineering", 228, "",
  "Biology", 208, ""
  )

kable(data, format = "latex", booktabs = TRUE,
      col.names = c(" ", "# Obtained", "Time to Completion"),
      align = c("l", "r", "l"),
      linesep = "") %>%
  kable_styling(full_width = TRUE) %>%
  pack_rows("Degree", 1, 5) %>%
  pack_rows("Majors", 6, 14) %>%
  column_spec(1, width = "7cm") %>%
  column_spec(2, width = "4.75cm") %>%
  row_spec(0, bold = TRUE,
           extra_css = "border-top: 0; border-bottom:5px solid red;") %>%
  add_header_above(c(" " , " ", " "))
```
\sdcoesource{National Student Clearinghouse StudentTracker Report.}
\sdcoenote{We calculated degree information and time to degree from the National Student Clearinghouse StudentTracker report using the Class of 2014. Many colleges and universities choose not to report degree titles or degree majors opting instead to report a graduation date only. For those students, we have identified their degree as "unreported."}

\newpage{}
\newgeometry{left=1in,right=1in,top=1in,bottom=1.5in}

\sdcoeheadingone{Final Thoughts}

Imagine you are one of the 40,648 graduating seniors from one of San Diego's 175 high schools. Going to college may have been an aspiration of yours for as long as you can remember. How do you feel right now? Perhaps you were already worried about how to pay for college, making friends, or failing a class. Now you may also be anxious about how the pandemic has changed your college dreams. What would it take to allay your fears and get you back on track for postsecondary options?

The fact is that Omicron or Covid-19 will have an impact on college-going for some time. We cannot end the pandemic as a solution to improving college access. The best chance we have of ending this enrollment crisis is to build stronger relationships with students and to provide them with information about their college-going options that is informative and keeps them progressing in the process. Additionally, we need to continue to understand the social and emotional stressors they experience in the process and  work to mitigate these challenges. Going to college is not something we can leave to chance. 

Everyone reading this report is an important decision-maker in his or her organization—many are directly responsible for college access. I urge you to stay informed about shifts in college enrollment patterns and invest in strategies that address students' fears and hesitations about enrolling in college. This group of high school graduates may be the most resilient of any other cohort before them. We must honor their experiences because they have truly paid a much greater price to achieve their aspirations than any of us every have or possibly will.


\sdcoeboilerplate{SDCOE's Assessment, Accountability, and Evaluation Department}{January}{2022}

