---
header-includes:
- \input{preamble.tex}
fontsize: 11pt
output: pdf_document
sansfont: Roboto
font:     Roboto
geometry: "left=1in,right=1in,top=0.35in,bottom=0.6in"
urlcolor: #169d62
---

\raggedright

```{r rmarkdown-setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
knitr::opts_chunk$set(message = FALSE)
knitr::opts_chunk$set(warning = FALSE)
knitr::opts_chunk$set(dev = "cairo_pdf")
options(knitr.kable.NA = "")
options(kableExtra.latex.load_packages = FALSE)

library(tidyverse)
library(knitr)
library(kableExtra)
library(sdcoethemes)

set_sdcoe_defaults(style = "print")
```

\sdcoelogo{}

\vspace{20mm}
\sdcoetitle{Declines in College Enrollment Continue Unabated in 2021}
\sdcoeauthors{Shannon Coulter, Director of Research and Evaluation}

\vspace{5mm}

\lettrine[lines=3, lraise=0.2, nindent=0em, slope=0em]{H}igher education enrollment fell to its lowest levels in San Diego in 2021. In San Diego County, an estimated 5,250 fewer students opted in to college compared to five years ago. Such a decline poses a serious risk for many students, especially those students who rely on a college degree to become upwardly mobile. And yet, simply enrolling in college is no panacea to upward mobility, students must actually attain a college degree, which is no small feat given that fewer than 25 percent actually do. Because of these historic declines in enrollment, we face a monumental problem: We currently have fewer students in the pipeline to attain a college degree--potentially losing 1,000 degree earners from the Class of 2021 alone--and this problem has the potential to further erode a fundamental purpose of K-12 education, our higher education institutions, and our economy at large.

Decisions about attending college were challenging prior to the pandemic. Students often questioned their readiness for the rigor of college work and the financial costs. The pandemic has layered other concerns that make college-going decisions even more unstable for students. 

There is no doubt that many students need a college degree to attain their professional aspirations and to achieve economic mobility. When someone's income improves over their lifetime that person is considered upwardly mobile, so those students completing a four-year degree have a better chance of becoming middle class than students who don't. Additionally, San Diego needs degree earners because we need to keep high-quality talent in the workforce in order to grow our economy. These problems are complex and we need to act on them in two ways. First, in the short-term, we must improve our understanding of college-going data through SDCOE's partnership with high school districts and the National Student Clearinghouse. That information will help educators address students' fears and hesitations about attaining a college degree and boost college access. Second, we must implement long-term solutions such as improving college and career readiness in the K-12 system and supporting lower income and disadvantaged students to complete college degrees.

The annual report of the National Clearinghouse data is designed to increase our understanding and decisions about college access and completion. The National Student Clearinghouse's StudentTracker service is a nationwide source of college enrollment and degree data from over 3,600 colleges and universities. The StudentTracker data provides our community with a reasonably accurate representation of the number and percentage of students who enroll, persist, and complete college annually. The SDCOE partnership with local high school districts to capture college-going data guide our efforts to support more students in attending and completing a four-year degree so they and our region can thrive.


\contactinfo{}

\newpage{}
\newgeometry{left=1in,right=1in,top=1in,bottom=1.5in}

\sdcoeheadingone{College Enrollment}

In 2021, the National Student Clearinghouse reported 18,915 San Diego County students enrolled in college right out of high school. College enrollment patterns in the Clearinghouse may vary from year to year for a variety of reasons including an overall decline in college going, fluctuations in the size of 12th grade cohorts, and the timing of when local high schools submit their graduates to the Clearinghouse. For that reason, this report is a preliminary report that is updated in late spring.

The map below locates the nearly 1,900 colleges and universities where students enrolled in 2021. Most students (83%) attended colleges and universities in Southern California, while the remaining 2,200 attended institutions in all other 49 states.

\sdcoefigurenumber{1}
\sdcoefiguretitle{While Many College Students Stay Close to Home, Graduates Enroll in Colleges and Universities Across the Country}

```{r enroll-map}
source("../nsc_21/scripts/09b usa_map.R", chdir = T)
```
\sdcoesource{National Student Clearinghouse StudentTracker Report}
\sdcoenote{We used geospatial data to identify the location of each college or university where graduates from the San Diego County Class of 2021 enrolled. We mapped each student to their respective college or university, darker dot patterns represent areas where more students attend college.}

\newpage{}
\newgeometry{left=1in,right=1in,top=1in,bottom=1.5in}

\sdcoeheadingtwo{Top 10 Colleges and Universities Attended}

For the first time in eight years more students enrolled in four-year institutions (9,633) compared to two-year (8,007). Some local community colleges suffered enrollment losses upward of 25% in 2021, while UC-San Diego reported a 26% increase in enrollment compared to the past two years. These variations in enrollment patterns have the greatest impact on lower-income students whose enrollment declines were double other student groups, according to the National Student Clearinghouse.

\vspace{3mm}
\sdcoetablenumber{1}
\sdcoetabletitle{More High School Graduates Opted for a Four-Year College Experience in 2021}
\sdcoetablesubtitle{Community colleges still have a big enrollment footprint in San Diego}

```{r echo=FALSE,  fig.width =7 , fig.height =2}
# arbitrary data to demonstrate a table
data <- tribble(~"College Name", ~Level, ~"Number Attending",
  "San Diego State University", "4-year", 1501,
  "University of California-San Diego", "4-year", 1310,
  "California State University-San Marcos", "4-year", 959,
  "University of California-Los Angeles", "4-year", 298,
  "Southwestern College", "2-year", 2004,
  "Palomar College", "2-year", 1827,
  "San Diego Mesa College", "2-year", 1073,
  "Grossmont College", "2-year", 953,
  "San Diego Miramar College", "2-year", 874,
  "San Diego City College", "2-year", 711,
  
  )

kable(data, format = "latex", booktabs = TRUE,
      col.names = c(" ", "Level", "Number Attending"),
      align = c("l", "r", "r"),
      linesep = "") %>%
  kable_styling(full_width = TRUE) %>%
  pack_rows("College Name", 1, 10) %>%
  column_spec(1, width = "7cm") %>%
  column_spec(2, width = "4.75cm") %>%
  row_spec(0, bold = TRUE,
           extra_css = "border-top: 0; border-bottom:5px solid red;") %>%
  add_header_above(c(" " , " ", " "))
```
\sdcoesource{National Student Clearinghouse StudentTracker Report}

\newpage{}
\newgeometry{left=1in,right=1in,top=1in,bottom=1.5in}

\sdcoeheadingtwo{College Enrollment by Cohort}

The overall number of students enrolled in college was down in 2021 compared to enrollment patterns prior to the pandemic. As indicated in the chart below, the number of students opting out of college (dark blue) increases over time, while the number enrolled (light blue) decreases over the same time period. Basically, a substantial number of students are opting out of college today compared to the past seven years.

\sdcoefigurenumber{2}
\sdcoefiguretitle{Early Enrollment Figures Show Another Steep Decline in College-Going}
\vspace{3mm}

```{r enroll-by-cohort}
source("../nsc_21/scripts/019a enroll_cohortv2.R", chdir = T)
```
\sdcoesource{National Student Clearinghouse StudentTracker Report}
\sdcoenote{We calculated enrollment number for each graduating cohort using data provided by the National Student Clearinghouse. The Clearinghouse data file supplies information on whether a student enrolled in college or not with a 93 percent matching rate.}

\newpage{}
\newgeometry{left=1in,right=1in,top=1in,bottom=1.5in}

\sdcoeheadingone{Persistence and Retention}

Despite the pandemic, students who enroll in college tend to stay. Some stay at their original institution (retention) while others opt to attend college elsewhere (persistence). While persistence and retention percentages declined slightly for the Class of 2020, differences between persistence and retention rates remained steady. About 4 in 5 enrolled students remained enrolled from year to year.

\sdcoefigurenumber{3}
\sdcoefiguretitle{Differences in College Persistence and Retention Hold Steady}
\sdcoefiguresubtitle{The number of students that remain in college compared to those staying in the same college}
\vspace{3mm}

```{r persist-by-cohort}
source("../nsc_21/scripts/021a persistence.R", chdir = T)
```
\sdcoesource{National Student Clearinghouse StudentTracker Report}
\sdcoenote{We calculated persistence as the percentage of students enrolled in college in the first year after high school graduation who then returned to any college for a second year. We calculated retention as the percentage of students returning to the same college from the previous year.}

\newpage{}
\newgeometry{left=1in,right=1in,top=1in,bottom=1.5in}

\sdcoeheadingone{Completion}

While the pandemic has made no discernible difference on college completion rates yet, historically only a  small number of students actually earn college degrees. About 19% of the Class of 2017 has earned a college degree, while 28% of the Class of 2015 earned a degree. To improve economic mobility, supporting students to access college isn't enough. We need to get students to finish. Furthermore, we might have significantly fewer students earning degrees in the future based on our current enrollment levels.

\sdcoefigurenumber{4}
\sdcoefiguretitle{More than 1 in 4 San Diego Students Earn a College Degree in 6 Years}
\sdcoefiguresubtitle{The percentange of students completing degrees in four and six years remains low}
```{r completion-by-cohort}
source("../nsc_21/scripts/012a completion_rates.R", chdir = T)
```
\sdcoesource{National Student Clearinghouse StudentTracker Report}
\sdcoenote{Colleges report graduation rates many different ways. We calculated graduation rates by cohort or year of high school graduation. The six-year graduation rate reflected on the graph represents the percentage of students from the Class of 2015 who earned a degree. The four-year rate represents only the students from the class of 2017 who have earned a degree.}

\newpage{}
\newgeometry{left=1in,right=1in,top=1in,bottom=1.5in}

\sdcoeheadingtwo{Degrees Earned}

For students who persisted and earned a degree, Business Administration topped the list as the most popular college major followed by Psychology and Computer Science. Because earnings vary based on the type of degree students get, several organizations including [College Scorecard](https://collegescorecard.ed.gov/) and [Salary Surfer](https://salarysurfer.cccco.edu/SalarySurfer.aspx) provide valuable information linking college degrees to the median salaries of graduates. Students may be surprised to learn that some associate's degrees may result in higher earning potential than some bachelor's degrees, which could reduce worries or hesitation about enrolling in a two- or four-year schools.

\sdcoetablenumber{2}
\sdcoetabletitle{2014 Graduates Earned Nearly 7,500 Degrees}
\sdcoetablesubtitle{Many students earn two-year degrees in same time some students earn four-year degrees}

```{r echo=FALSE,  fig.width =7 , fig.height =2}
# arbitrary data to demonstrate a table
data <- tribble(~"Degree", ~"# Obtained", ~"Time to Completion",
  "Certificate", 131, "3 years, 11 months",
  "Associate's", 347, "4 years 1 month",
  "Bachelor's", 3082, "4 years 6 months",
  "Master's and beyond (with Bachelor's)", 352, "5 years 11 months",
  "Unreported", 3418, "4 years 6 months",
  "Business Administration", 722, "",
  "Psychology", 503, "",
  "Computer Science", 424, "",
  "Political Science", 256, "",
  "Economics", 254, "",
  "Sociology", 250, "",
  "Communication", 229, "",
  "Mechanical Engineering", 228, "",
  "Biology", 208, ""
  )

kable(data, format = "latex", booktabs = TRUE,
      col.names = c(" ", "# Obtained", "Time to Completion"),
      align = c("l", "r", "l"),
      linesep = "") %>%
  kable_styling(full_width = TRUE) %>%
  pack_rows("Degree", 1, 5) %>%
  pack_rows("Majors", 6, 14) %>%
  column_spec(1, width = "7cm") %>%
  column_spec(2, width = "4.75cm") %>%
  row_spec(0, bold = TRUE,
           extra_css = "border-top: 0; border-bottom:5px solid red;") %>%
  add_header_above(c(" " , " ", " "))
```
\sdcoesource{National Student Clearinghouse StudentTracker Report}
\sdcoenote{We calculated degree information and time to degree from the National Student Clearinghouse StudentTracker report using the Class of 2014. Many colleges and universities choose not to report degree titles or degree majors opting instead to report a graduation date only. For those students, we have identified their degree as "unreported."}

\newpage{}
\newgeometry{left=1in,right=1in,top=1in,bottom=1.5in}

\sdcoeheadingone{Final Thoughts}

Imagine you are one of the 40,648 graduating seniors from one of San Diego's 175 high schools. Going to college may have been an aspiration of yours for as long as you can remember. How do you feel right now? Perhaps you were already worried about how to pay for college, making friends, or failing a class. Now you may also be anxious about how the pandemic has changed your college dreams. What would it take to allay your fears and get you back on track for postsecondary options?

The fact is that Covid-19 or any of its variants will have an impact on college-going for some time. We cannot end the pandemic as a solution to improving college access. The best chance we have of ending this enrollment crisis is to build stronger relationships with students and to provide them with information about their college-going options that keeps them progressing. Additionally, we need to better understand the social and emotional stressors they experience in the process and  work to mitigate these challenges. Going to college is not something we can leave to chance. 

Everyone reading this report is an important decision-maker in their organization—many are directly responsible for college access. I urge you to stay informed about shifts in college enrollment patterns and invest in strategies that address students' fears and hesitations about enrolling in college. This group of high school graduates may be the most resilient of any other cohort before them. We must honor their experiences because they have truly paid a much greater price to achieve their aspirations than any of us ever have or possibly will.

\vspace{10mm}
\begin{tcolorbox}[colback=sdcoelightblue!5, colframe=sdcoelightblue!75]
\sdcoeheadingone{Interested in boosting four-year college enrollment rates?}

Check out these resources:

\begin{sdcoebullets}
  \item \href{https://resources.finalsite.net/images/v1633543351/sdcoenet/n9ck4tbjbsexckrvxrgb/CollegeAccess-Brochure.pdf}{Is a four-year college right for me?}
  \item \href{https://resources.finalsite.net/images/v1639429972/sdcoenet/kmq4ljy195amwtiinc4w/MakeItEasier-Report.pdf}{Make it Easier}
  \item \href{https://www.sdcoe.net/educators/evaluation/improvement-networks#fs-panel-2947}{College Enrollment Checklists}
\end{sdcoebullets}
\end{tcolorbox}

\sdcoeboilerplate{SDCOE's Assessment, Accountability, and Evaluation Department}{February}{2022}

